---
title: "Map ROMS data to Ecopath GOA"
author: "Alberto Rovellini"
date: "5/11/2021"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

Based on code from Owen Liu and Alberto Rovellini.

# Purpose

This document illustrates the workflow to pull static variables from ROMS NetCDF output and map it to the model area of Ecopath models for the Gulf of Alaska. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,include=FALSE}
library(tidyverse)
library(tidync)
library(sf)
library(rnaturalearth)
library(raster)
library(data.table)
library(maps)
library(mapdata)
library(angstroms)
library(viridis)
```

```{r}
select <- dplyr::select
```

# Read data

## Read Ecopath mask

Load the Ecopath spatial masks.
```{r}
mask <- st_read("EGOA.shp")
#mask <- mask %>% st_combine()
```

Have a look.
```{r}
coast <- map("worldHires", regions = "USA", plot = FALSE, fill = TRUE)
coast <- coast %>% st_as_sf() %>% st_transform(crs = st_crs(mask)) %>% st_crop(st_bbox(mask))

ggplot()+
  geom_sf(data = mask, fill = "coral")+
  geom_sf(data = coast, fill = "grey")+
  theme_minimal()+
  labs(title = "EGOA Ecopath mask")
```

## Import ROMS data

For GOA, we have grid information stored in a grid file, and the variables stored in the netCDF files.

```{r}
romsfile <- "data/ROMS/test_from_loon/cgoa_moave_2016_04.nc"
romsfile2 <- "data/ROMS/CGOA_grid_5.nc"
```

```{r}
roms_vars <- tidync(romsfile)
roms_grid <- tidync(romsfile2)
```

Some information about this NetCDF.
```{r}
print(roms_vars)
```

Get variables. We do not need water velocity, so we can ignore $u$ and $v$ points. Also we should not need depth, so actually all that matters is 2D indexing for $\rho$ points.
```{r}
#grid info
grid_variables <- hyper_grids(roms_grid) %>% # all available grids in the ROMS ncdf
  pluck("grid") %>% # for each grid, pull out all the variables asssociated with that grid and make a reference table
  purrr::map_df(function(x){
    roms_grid %>% activate(x) %>% hyper_vars() %>% 
      mutate(grd=x)
  })

#variables
roms_variables <- hyper_grids(roms_vars) %>% # all available grids in the ROMS ncdf
  pluck("grid") %>% # for each grid, pull out all the variables asssociated with that grid and make a reference table
  purrr::map_df(function(x){
    roms_vars %>% activate(x) %>% hyper_vars() %>% 
      mutate(grd=x)
  })
```

Get some information about time.
```{r}
ncmeta::nc_atts("data/ROMS/three_steps.nc", "ocean_time") %>% tidyr::unnest(cols = c(value))

time_grd <- roms_variables %>% filter(name=="ocean_time") %>% pluck('grd')
# pull the lon/lats
roms_time <- roms_vars %>% activate(time_grd) %>% hyper_tibble() %>% pull()
```

```{r}
# find appropriate ROMS ncdf grid for the rho points
latlon_rhogrd <- grid_variables %>% filter(name=="lat_rho") %>% pluck('grd')
# pull the lon/lats
roms_rho <- roms_grid %>% activate(latlon_rhogrd) %>% hyper_tibble() %>% select(lon_rho,lat_rho,xi_rho,eta_rho) %>% 
  mutate(rhoidx=row_number()) # add index
```

Add coordinates in the CRS used by the Ecopath mask.
```{r}
append_xy_coords <- function(lonlatdat, xyproj=crs(mask)@projargs, lon_col="lon_rho", lat_col="lat_rho"){
  lonlatdat %>% 
    st_as_sf(coords=c(lon_col, lat_col), crs=4326, remove=F) %>%  # convert to spatial object
    st_transform(xyproj) %>%  # convert to Atlantis coords
    mutate(x = st_coordinates(.)[,1],
           y = st_coordinates(.)[,2]) # grab x and y coordinates and add them as attributes
}

rhoxy<- append_xy_coords(roms_rho,lon_col="lon_rho",lat_col="lat_rho") %>% mutate(rhoidx=row_number())
```

# Match ROMS and Ecopath

Join $\rho$ points to the Ecopath mask. Do a spatial join.
```{r}
rho_join <- mask %>% st_join(rhoxy) %>% na.omit()
```

Test that we are pulling the right points.
```{r}
rho_join %>% 
  st_set_geometry(NULL) %>% 
  st_as_sf(coords = c("x", "y")) %>%
  ggplot()+
  geom_sf(size = 0.1)+
  theme_minimal()+
  labs(title = "CGOA 3K rho points overlapping with Ecopath mask")
```

Get indeces of rho, u, and v points that overlap with Atlantis geometry, to subset large ROMS files and reduce memory chokes
```{r}
min_xi_rho <- min(rho_join$xi_rho, na.rm = TRUE)
max_xi_rho <- max(rho_join$xi_rho, na.rm = TRUE)
min_eta_rho <- min(rho_join$eta_rho, na.rm = TRUE)
max_eta_rho <- max(rho_join$eta_rho, na.rm = TRUE)
```

# Set up depths

Using a custom version of Mike Sumner's angstroms::romshcoords(), because GOA ROMS have grid information in a separate file.
```{r}
ncget <- function(x, varname) {
  nc <- ncdf4::nc_open(x)
  on.exit(ncdf4::nc_close(nc))
  ncdf4::ncvar_get(nc, varname)
}

set_indextent <- function(x) {
  setExtent(x, extent(0, ncol(x), 0, nrow(x)))
}

romshcoords_goa <- function(x, y, grid_type = "rho", slice, ..., S = "Cs_r", depth = "h", simple = FALSE){
  h <- romsdata(x, varname = depth)
  Cs_r <- ncget(y, S)
  v <- values(h)
  if (simple) {
    ## simplistic, early version - probably should be defunct
    out <- set_indextent(brick(array(rep(rev(Cs_r), each = length(v)) * v, 
                                     c(ncol(h), nrow(h), length(Cs_r))), transpose = TRUE))
  } else {
    grid_type <- match.arg(tolower(grid_type),c("rho","psi","u","v","w"))
    
    Vtransform <- as.integer(ncget(y,"Vtransform"))
    if (!Vtransform %in% c(1,2)) stop("Vtransform must be 1 or 2")
    
    hc <- ncget(y,"hc")
    
    depth_grid <- if (grid_type=="w") "w" else "rho"
    
    zeta <- if (missing(slice)) 0 else stop("slice not supported yet")##angstroms::romsdata2d(x,"zeta",slice=slice,transpose=FALSE)
    N <- length(ncget(y,"Cs_r"))
    Np <- N+1
    
    h <- ncget(x,"h")
    hmin <- min(h)
    hmax <- max(h)
    
    Lp <- dim(h)[1]
    Mp <- dim(h)[2]
    L <- Lp-1
    M <- Mp-1
    
    z <- array(NA,dim=c(Lp,Mp,if (grid_type=="w") Np else N))
    
    ## Compute vertical stretching function, C(k):
    ##stretch <- stretching(x,depth_grid)
    if (depth_grid=="w") {
      stretch <- list(C=ncget(y,"Cs_w"),s=ncget(y,"s_w"))
    } else {
      stretch <- list(C=ncget(y,"Cs_r"),s=ncget(y,"s_rho"))
    }
    
    ## Average bathymetry and free-surface at requested C-grid type.
    if (grid_type=="rho") {
      hr <- h
      zetar <- zeta
    } else if (grid_type=="psi") {
      hp <- 0.25*(h[1:L,1:M]+h[2:Lp,1:M]+h[1:L,2:Mp]+h[2:Lp,2:Mp])
      zetap <- 0.25*(zeta[1:L,1:M]+zeta[2:Lp,1:M]+zeta[1:L,2:Mp]+zeta[2:Lp,2:Mp])
    } else if (grid_type=="u") {
      hu <- 0.5*(h[1:L,1:Mp]+h[2:Lp,1:Mp])
      zetau <- 0.5*(zeta[1:L,1:Mp]+zeta[2:Lp,1:Mp])
    } else if (grid_type=="v") {
      hv <- 0.5*(h[1:Lp,1:M]+h[1:Lp,2:Mp])
      zetav <- 0.5*(zeta[1:Lp,1:M]+zeta[1:Lp,2:Mp])
    } else if (grid_type=="w") {
      hr <- h
      zetar <- zeta
    } else {
      stop("unsupported grid_type: ",grid_type)
    }
    
    ## Compute depths (m) at requested C-grid location.
    
    if (Vtransform == 1) {
      if (grid_type=="rho") {
        for (k in seq_len(N)) {
          z0 <- (stretch$s[k]-stretch$C[k])*hc + stretch$C[k]*hr
          z[,,k] <- z0 + zetar*(1.0 + z0/hr)
        }
      } else if (grid_type=="psi") {
        for (k in seq_len(N)) {
          z0 <- (stretch$s[k]-stretch$C[k])*hc + stretch$C[k]*hp
          z[,,k] <- z0 + zetap*(1.0 + z0/hp)
        }
      } else if (grid_type=="u") {
        for (k in seq_len(N)) {
          z0 <- (stretch$s[k]-stretch$C[k])*hc + stretch$C[k]*hu
          z[,,k] <- z0 + zetau*(1.0 + z0/hu)
        }
      } else if (grid_type=="v") {
        for (k in seq_len(N)) {
          z0 <- (stretch$s[k]-stretch$C[k])*hc + stretch$C[k]*hv;
          z[,,k] <- z0 + zetav*(1.0 + z0/hv)
        }
      } else if (grid_type=="w") {
        z[,,1] <- -hr
        for (k in seq(from=2,to=Np,by=1)) {
          z0 <- (stretch$s[k]-stretch$C[k])*hc + stretch$C[k]*hr
          z[,,k] <- z0 + zetar*(1.0 + z0/hr)
        }
      } else {
        stop("unsupported grid_type: ",grid_type)
      }
    } else if (Vtransform == 2) {
      if (grid_type=="rho") {
        for (k in seq_len(N)) {
          z0 <- (hc*stretch$s[k]+stretch$C[k]*hr)/(hc+hr)
          z[,,k] <- zetar+(zeta+hr)*z0
        }
      } else if (grid_type=="psi") {
        for (k in seq_len(N)) {
          z0 <- (hc*stretch$s[k]+stretch$C[k]*hp)/(hc+hp)
          z[,,k] <- zetap+(zetap+hp)*z0
        }
      } else if (grid_type=="u") {
        for (k in seq_len(N)) {
          z0 <- (hc*stretch$s[k]+stretch$C[k]*hu)/(hc+hu)
          z[,,k] <- zetau+(zetau+hu)*z0
        }
      } else if (grid_type=="v") {
        for (k in seq_len(N)) {
          z0 <- (hc*stretch$s[k]+stretch$C[k]*hv)/(hc+hv)
          z[,,k] <- zetav+(zetav+hv)*z0
        }
      } else if (grid_type=="w") {
        for (k in seq_len(Np)) {
          z0 <- (hc*stretch$s[k]+stretch$C[k]*hr)/(hc+hr)
          z[,,k] <- zetar+(zetar+hr)*z0
        }
      } else {
        stop("unsupported grid_type: ",grid_type)
      }
    } else {
      stop("Vtransform must be 1 or 2")
    }
    ## FIXME all these flips and twirls can be applied more efficiently (or avoided)
    ## though should layers start at the surface and go down or ...
    
    out <- raster::flip(set_indextent(raster::brick(z, transpose = TRUE)), "y")
    ## NO - we want to start at the bottom, so we match romsdata3d
    #out <- raster::subset(out, rev(seq_len(raster::nlayers(out))))
    
  } 
  
  out
}
```

Apply to get depths at $\rho$ points. For the purpose of the Ecopath model this will only be done once, altough depth in ROMS is dynamic over time. We only do it once in Atlantis too - for models like these the bending free surface amounts to rounding error.
```{r}
# convert ROMS s-coordinates to depth with Mike Sumner's angstroms package

romsdepths <- romshcoords_goa(x = romsfile2, y = romsfile, S = "Cs_r", depth = "h")

# using tabularaster to convert to tibble
# and a indexing template with "by_column" filling
romsi <- crossing(xi_rho=1:dim(romsdepths)[2],eta_rho=1:dim(romsdepths)[1]) %>% arrange(-eta_rho) %>% mutate(cellindex=row_number()) # making sure that the join by cellindex below is correct - doing this for consistency with the way tabularaster::as_tibble() unpacks the raster cells 
romsdepthsdf <- tabularaster::as_tibble(romsdepths,dim=F) %>% 
  arrange(cellindex) %>% 
  left_join(romsi,by='cellindex') %>% 
  set_names(c("romsdepth","cellindex","xi_rho","eta_rho")) %>% 
  group_by(cellindex,xi_rho,eta_rho) %>% 
  nest(romsdepth=c(romsdepth)) %>% ungroup() %>% 
  mutate(romsdepth=purrr::map(romsdepth,function(x)x[['romsdepth']])) %>%
  filter(between(xi_rho, min_xi_rho, max_xi_rho) & between(eta_rho, min_eta_rho, max_eta_rho))
```

# Pull variables from ROMS

Writing a function to interpolate variables with cubic splines over 1 m intervals in the water column. 
```{r}
interp_foo <- function(romsdepths,romsvar) {
  depths_out <- seq(round(min(romsdepths)),0,by=1) # 1m interpolation, starting from deepest
  interp <- spline(romsdepths,romsvar,xout=depths_out) %>% pluck('y')
  return(tibble(depth=depths_out,val=interp))
}
```

Writing another function that:

* Applies the cubic spline interpolation at each $\rho$ point.
* Integrates over the water column as appropriate for each variable (i.e. average for state variables like temperature and salinity, sum for concentrations per m$^3$ to obtain values per m$^2$).
* Return average values for each traces over the Ecopath model domain.
```{r}
interpolate_var <- function(variable, time_step){
  grd <- roms_variables %>% filter(name==variable) %>% pluck('grd')
  # pull the env data
  # interpolate the env data
  # do this step conditional to join with the appropriate depth data frame depending on the variable
  # if variable is horizontal velocity
  dat <- roms_vars %>% activate(grd) %>%
    hyper_tibble(select_var=variable, 
                 xi_rho = between(xi_rho, min_xi_rho, max_xi_rho), 
                 eta_rho = between(eta_rho, min_eta_rho, max_eta_rho),
                 ocean_time = ocean_time == time_step)
  
  interp_dat <- dat %>% 
    dplyr::select(xi_rho,eta_rho,!!variable) %>% 
    nest(data=c(!!variable))%>% 
    mutate(evar=purrr::map(data,~.[[1]]))
  
  interp_dat <- interp_dat %>%
    inner_join(romsdepthsdf,by=c('xi_rho','eta_rho')) 
  
  interp_dat <- interp_dat %>% 
    mutate(interp = purrr::map2(romsdepth,evar,interp_foo)) %>% 
    dplyr::select(-data,-evar,-romsdepth)
  
  #TODO: spit a warning if for any rho point temp at the surface is lower than at depth - may be sign of depths from ROMS being inverted
  
  # integrate. For state variables, take the average over the water column. For concentrations, sum over the water column.
  if(variable %in% c(state_vars)){
    interp_dat <- interp_dat %>%
    mutate(value_m2 = purrr::map_dbl(interp,function(x)mean(x$val)),
           depth = purrr::map_dbl(interp,function(x)min(x$depth))) %>%
      select(-interp)
  } else {
    interp_dat <- interp_dat %>%
    mutate(value_m2 = purrr::map_dbl(interp,function(x)sum(x$val)),
           depth = purrr::map_dbl(interp,function(x)min(x$depth))) %>%
      select(-interp)
  }
  
  # join to rho_join set to subset to Ecopath mask only
  
  interp_dat <- rho_join %>%
    st_set_geometry(NULL) %>%
    left_join(interp_dat,by=c('xi_rho','eta_rho'))
  
  # return(interp_dat) # if you want to see what things look like spatially, this is a good place to return an output for visualisation
  
  # take averages over the entire model domain
  ecopath_value <- interp_dat %>% 
    select(value_m2) %>% 
    summarise(ecopath_value=mean(value_m2,na.rm=TRUE)) %>%
    ungroup() %>%
    pull()
}
```

Apply to all variables in all time steps.

Enumerate the variables of interest.
```{r}
state_vars <- c('temp','salt','frat_PhS','frat_PhL','CChl_PhS','CChl_PhL')
conc_vars <- c('NO3','NH4','PhS','PhL','MZS','MZL','Cop','NCa','Eup','Det','Iron','prod_PhS','prod_PhL','prod_MZS','prod_MZL','prod_Cop','prod_NCa','prod_Eup')
```

```{r}
start<- Sys.time()

all_variables <- c(state_vars,conc_vars)
# roms_time <- roms_time[1:2]

var_time_combos <- expand.grid(all_variables,roms_time) %>% mutate(Var1=as.character(Var1)) %>%
  set_names(c('variable','time_step'))

ecopath_vals <- var_time_combos %>% 
  mutate(ecopath_values = purrr::pmap_dbl(list(variable=variable,time_step=time_step),interpolate_var))

epoch <- "1900-01-01" #important, check that this is your correct start - keeping this generic as it changes a lot ROMS by ROMS

ecopath_vals <- ecopath_vals %>% mutate(date=as.POSIXct(time_step, origin = epoch, tz='UTC')) %>%
  arrange(variable,time_step)

stop <- Sys.time()
stop-start
```

View.
```{r}
ecopath_vals %>%
  ggplot(aes(x = date, y = unlist(ecopath_values)))+
  geom_point()+
  geom_line()+
  theme_minimal()+
  facet_wrap(~variable, scales = "free")+
  xlab("Day")+
  ylab("Value")
```
